---
title: "Red Wine Quality"
author: "Jin Seo Jo"
date: "26/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(scales)
```

```{r}
wine_data <- read.csv("winequality-red.csv")
```

## Wine Quality
```{r}
# Distribution of red wine quality ratings
ggplot(wine_data, aes(x = quality)) +
  geom_bar(stat = "count", position = "dodge", fill = "red") +
  ggtitle("Distribution of Red Wine Quality Ratings")
```

## Distribution of good/bad wines
We will classify the wines by setting an arbitrary cutoff for our dependent variable (wine quality) at e.g. 7 or higher getting classified as 'good/1' and the remainder as 'not good/0'.
```{r}
# Create a variable indicating if a wine is good or bad
wine_data$good_bad_wine <- ifelse(wine_data$quality > 6, "Good", "Bad") %>% 
  factor()

summary(wine_data)
```
`good_bad_wine` tells us that only 217 out of 1599 wines is considered as a good wine
```{r}
# Distribution of good/bad wines
ggplot(wine_data, aes(x = factor(1), fill = good_bad_wine)) +
  geom_bar(width = 1) +
  coord_polar("y") +
  labs(fill = "Wine quality") +
  theme_void()
```
 
## Relationship between Physiochemical Properties and Wine Quality
Input variables (based on physiochemical tests):
```{r}
ls(wine_data)
```

### 1. Fixed Acidity and Wine Quality
```{r}
# From the 'wine_data', 
# select rows where 'good_bad_wine' is "Good" and recode column 'fixed.acidity'
wine_data$fixed.acidity[wine_data$good_bad_wine == "Good"]

# Find mean values
mean(wine_data$fixed.acidity[wine_data$good_bad_wine == "Good"])
mean(wine_data$fixed.acidity[wine_data$good_bad_wine == "Bad"])
```

```{r}
ggplot(wine_data, aes(x = fixed.acidity, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  # 'na.rm = TRUE' will remove NA values before the computation proceeds
  geom_vline(aes(xintercept = mean(fixed.acidity[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(fixed.acidity[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Fixed Acidity Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Fixed Acidity Levels") +
  theme_classic()
```

### 2. Volatile Acidity and Wine Quality
```{r}
ggplot(wine_data, aes(x = volatile.acidity, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(volatile.acidity[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(volatile.acidity[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Volatile Acidity Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Volatile Acidity Levels") +
  theme_classic()
```
 
 ### 3. Citric Acid and Wine Quality
```{r}
ggplot(wine_data, aes(x = citric.acid, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(citric.acid[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(citric.acid[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Citric Acidity Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Citric Acidity Levels") +
  theme_classic()
```
 
 ### 4. Residual Sugar and Wine Quality
```{r}
ggplot(wine_data, aes(x = residual.sugar, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(residual.sugar[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(residual.sugar[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Residual Sugar Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Residual Sugar Levels") +
  theme_classic()
```
 
Good and bad wines have very similar distributino of their corresponding physiochemical properties. 
 
### 5. Chlorides and Wine Quality
```{r}
ggplot(wine_data, aes(x = chlorides, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(chlorides[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(chlorides[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Chlorides Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Chlorides Levels") +
  theme_classic()
```

Good and bad wines have very similar distributino of their corresponding physiochemical properties. 

### 6. Free Sulfur Dioxide and Wine Quality
```{r}
ggplot(wine_data, aes(x = free.sulfur.dioxide, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(free.sulfur.dioxide[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(free.sulfur.dioxide[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Free Sulfur Dioxide Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Free Sulfur Dioxide Levels") +
  theme_classic()
```

Good and bad wines have very similar distributino of their corresponding physiochemical properties. 

### 7. Total Sulfur Dioxide and Wine Quality
```{r}
ggplot(wine_data, aes(x = total.sulfur.dioxide, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(total.sulfur.dioxide[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(total.sulfur.dioxide[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Total Sulfur Dioxide Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Total Sulfur Dioxide Levels") +
  theme_classic()
```

Good and bad wines have very similar distributino of their corresponding physiochemical properties. 

### 8. Density and Wine Quality
```{r}
ggplot(wine_data, aes(x = density, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(density[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(density[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Density Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Density Levels") +
  theme_classic()
```

 Good and bad wines have very similar distributino of their corresponding physiochemical properties. 

### 9. pH and Wine Quality
```{r}
ggplot(wine_data, aes(x = pH, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(pH[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(pH[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "pH Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of pH Levels") +
  theme_classic()
```

 Good and bad wines have very similar distributino of their corresponding physiochemical properties. 

### 10. Sulphates and Wine Quality
```{r}
ggplot(wine_data, aes(x = sulphates, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(sulphates[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(sulphates[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Sulphates Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Sulphates Levels") +
  theme_classic()
```

Sulphates level of the wine shows the most discriminating attributes.

### 11. Alcohol and Wine Quality
```{r}
ggplot(wine_data, aes(x = alcohol, fill = good_bad_wine)) + 
  geom_density(alpha = 0.25) +
  geom_vline(aes(xintercept = mean(alcohol[good_bad_wine == "Good"], na.rm = TRUE)), 
             color = "red", linetype = "dashed", lwd = 1) +
  geom_vline(aes(xintercept = mean(alcohol[good_bad_wine == "Bad"], na.rm = TRUE)), 
             color = "blue", linetype = "dashed", lwd = 1) +
  scale_x_continuous(breaks = seq(4, 16, 1)) +
  xlab(label = "Alcohol Level") +
  labs(fill = "Wine Quality") +
  ggtitle("Distribution of Alcohol Levels") +
  theme_classic()
```

Alchol level of the wine shows the most discriminating attributes.

**Interpretation**:  
- We can see that `residual.sugar`, `chlorides`, `free.sulfur.dioxide`, and `pH` have similar distributions. This means that `residual.sugar`, `chlorides`, `free.sulfur.dioxide`, and `pH` does not effect on wine quality.  
- The most discriminating attributes we can observe are `sulphates` and `alcohol` level of the wine. This shows that `sulphates` and `alcohol` improve the quality of wine.

## Find the most important variables
Selecting the most important predictor variables that explains the major part of variance of the reponse variable can be key to identify and build high performing models.

### Random Forest Method
Random forest can be very effectve to find a set of predictors that best explains the variance in the reponse variable.

Random forest are made out of decision trees. 
Decision Trees are easy to build, easy to use and easy to interpret, but in practice they are not that awesome. 
Decision Trees work great with the data used to create them, but they are not flecible when it comes to classifying new samples. 
Random Foressts combine the simplicity of decision trees with flexibility resulting in a vast improvement in accuracy.


```{r, message=FALSE}
library(caret)
library(randomForest)
library(varImp)
```

```{r}
str(wine_data)
```

```{r}
sum(is.na(wine_data))
```

We do not have any missing values, therefore we do not have to impute values for the NAs in the dataset with 'rfImput()'

```{r}
# Fit the random forest
# We will remove 'quality' since 'good_bad_wine' is based on 'quality'
# We want the good_bad_wine column to be predicted by the data in all of the other columns except quality column
model <- randomForest(formula = good_bad_wine ~ . - quality, data = wine_data, proximity = TRUE)

model
```

**Type of random forest** show athat the random forest was built to classify samples.

**Number of trees** tells us how many trees are in the random forest. 
Note: The default value is 500.

**No. of variables tried at each split** tells us how many variables (or columns of data) were considered at each internal node. 
Note: Classification trees have a default setting of the square root of the number of variables. Regression trees have a default setting of the number of variables divided by 3. 

**OOB estimate of error rate** means that 92% of OOB(Out-of-Bag) samples were correctly classified by the random forest. 

**Confusion matrix**:
- There were 1348 bad quality wines that were correctly labeled "Bad"
- There were 98 good quality wines that were incorrectly classified "Bad"
- There were 34 bad quality wines that were incorrectly classified "Good"
- There were 119 good quality wines that were correctly classified "Good"

To see if 500 trees is enough for optimal classification, we can plot the error rates.
```{r}
oob.error.data <- data.frame(
  Trees = rep(1:nrow(model$err.rate), times = 3),
  Type = rep(c("OOB", "Bad", "Good"), each = nrow(model$err.rate)),
  Error = c(model$err.rate[,"OOB"],
            model$err.rate[, "Bad"],
            model$err.rate[, "Good"]))
```

For the most part, this is all based on a matrix within `model` called `err.rate`

More detail on `err.rate`:
```{r}
head(model$err.rate)
```

- There is one column for OOB error rate
- There is one column for the Bad error rate (i.e. how frequently bad wines are misclassified)
- There is one column for the Good error rate (i.e. how frequently good wines are misclassified)
- Each row reflects the error rates at different stages of creating the random forest.
- The first row contains the error rates after making the first tree.
- The second row contains the error rates after making the first two trees.
- The last row contians the error rates after making all 500 trees (i.e. default value)

More detail on `oob.error.data`:
```{r}
head(oob.error.data)
```

- There is one column for the number of trees.
- There is one column for the type of error.
- There is one column for the actual error value.

Now we call `ggplot()`
```{r}
ggplot(data = oob.error.data, aes(x = Trees, y = Error)) +
  geom_line(aes(color = Type))
```

- The green line shows the error rate when classifying good quality of wines. 
- The blue line shows the overall OOB error rate.
- The red line shows the error rate when classifying bad quality of wines.
- In general, we see the error rates decrease when our random forest has more trees. 
- And we see that the error rates stabilize right after 200 trees. 
- So adding more trees didn't help, but we would not have known this unless we used more trees. 

Now we need to make sure we are considering the optimal number of variables at each internal node in the tree.  
We start by making an empty vector that can hold 10 values.
```{r}
oob.values <- vector(length = 10)
```
Then we create a loop that tests differnet numbers of variables at each step. 
```{r}
# Each time we go through the loop, "i" increases by 1.
# It starts at 1 and ends after 10.
for(i in 1:10) {
  # We are building a random forest using "i" to determine the numebr of variables to try at each step
  # Specifically, we are setting 'mtry = i', and "i" equals values between 1 and 10
  temp.model <- randomForest(formula = good_bad_wine ~ . - quality, data = wine_data, mtry = i)
  # We store the OOB error rate after we build each random forest that uses a different value for mtry
  # temp.model contains a matrix called err.rate
  # And we want to access the value in the last row and in the first column
  # (i.e. the OOB error rate when all 500 trees have been made)
  oob.values[i] <- temp.model$err.rate[nrow(temp.model$err.rate), 1]
}
```
Print out the OOB error rates for different values for `mtry`
```{r}
oob.values
```
- The 3rd value, corresponding to `mtry = 3`, which is the default in this case, does not have the lowest OOB error rate.  
- So the default value was not optimal.  
- We need to choose the 2nd value since it has the lowest OOB error rate.  

Make a new random forest model:
```{r}
new_model <- randomForest(formula = good_bad_wine ~ . - quality, data = wine_data, ntree = 200, mtry = 2, proximity = TRUE)

new_model
```

Lastly, we want to use the random forest to draw an MDS plot with samples.  
This will show us how they are related to each other.  

We start by using `dist()` to make a distance matrix from 1-proximity matrix
```{r}
distance.matrix <- dist(1 - new_model$proximity)
```
Then we run `cmdscale()` on the distance matrix.
```{r}
# Classical Multidimensional Scaling
mds.stuff <- cmdscale(distance.matrix, eig = TRUE, x.ret = TRUE)
```
Then we calculate the percentage of variation in the distance matrix that the X and Y axes account for. 
```{r}
mds.var.per <- round(mds.stuff$eig/sum(mds.stuff$eig)*100, 1)
```
Then we format the data for `ggplot()`
```{r}
mds.values <- mds.stuff$points

mds.data <- data.frame(Sample = rownames(mds.values),
                       X = mds.values[,1],
                       Y = mds.values[,2],
                       Status = wine_data$good_bad_wine)
```
Draw the graph with `ggplot()`
```{r}
ggplot(data = mds.data, aes(x = X, y = Y, label = Sample)) +
  geom_text(aes(color = Status)) +
  xlab(paste("MDS1 - ", mds.var.per[1], "%", sep = "")) +
  ylab(paste("MDS2 - ", mds.var.per[2], "%", sep = "")) +
  ggtitle("MDS plot using (1 - Random Forest Proximities)")
```
- We can see that Good sampels are on the right side.  
- The X-axis accounts for 52.7% of the variation in the distance matrix.  
- The y-axis only accounts for 8.9% of the variation in the distnace matrix.  
- That means that the big differences are along the X-axis.


